A performance comparison of Dask and Spark for data-intensive neuroimaging pipelines

### 2019-07-26 latex/paper.tex
Modified results section + Positioned the figures

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{times}
\usepackage[binary-units=true]{siunitx}
\usepackage{latexsym}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}

@@

While Python multithreading is limited by the
\href{https://docs.python.org/3/glossary.html#term-gil}{Global Interpreter Lock (GIL)}
as a result of CPython's non thread-safe memory management, Dask attempts to
bypass the GIL whenever possible and use multithreading to provide parallelism.
With the exception of Dask Bags, all APIs, by default, use the local multithreaded
scheduler. Dask Bags instead relies on the local multiprocessing scheduler.

@@

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.125\textwidth,
    angle=-90]{images/incrementation-task-graph.png}
    \caption{Task graph for Incrementation with 5 iterations and 3 BigBrain blocks.
    Circles represent the incrementation and sleep function while rectangles
    represent stages of the image blocks.}\label{fig:tg-inc}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.16\textwidth, angle=-90]{images/histogram-task-graph.png}
    \caption{Task graph for Histogram with 3 BigBrain blocks. Circles represent the
    functions while rectangles represent stages of the image blocks.
    }\label{fig:tg-histo}
\end{figure}

@@

While Incrementation and Histogram were implemented directly in Python, BIDS app
example requires an external library and is distributed as a Docker container image
(bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image,
given its popularity for HPC and reproducibility, using
\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}.
The image was preloaded on the nfs and singularity version 3.2.1\-1.el7 was installed
on all instances. Due to a bug in the conversion from Docker to Singularity, we had to
export the path for fsl.

@@

For each run, we measured the application makespan as well as the cumulative 
data read, compute, data write, and engine overhead across all application tasks. 
We measured the overhead by \TG{\ldots, maybe add a figure}. \MD{Done but might need
adjustment depending on how the overhead is calculate (idle vs wasted)}

\TG{explain that number of iterations may impact futures.}

% The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
% seconds sleep delay.

@@

%%% INCREMENTATION %%%
%% worker
\subsection{Incrementation: Number of workers}
Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation application
for different numbers of workers and for different engines. The bars show the average
makespan over 3 repetitions while the error bars are the standard deviation. Overall,
there is no substantial makespan difference between the engines. Dask seems to have a
slight advantage over Spark, \SI{83.61}{\second} on average,
with Delayed and Futures being slightly better than Bags.

For all engines, the makespan is far from decreasing linearly with the
number of workers. Note that there are 8 threads per worker. The makespan
even increases between 4 and 8 workers. This is due to the high impact of
data transfers and engine overhead on the application.

Figure~\ref{fig:inc_tt_worker} shows the total execution time of the
Incrementation application, broken down into data transfer (read and
write), compute (sleep), and overhead time. As expected, the computing time
stays similar when the number of workers increases. However, the data
transfer time and overhead increase proportionally to the number of
workers \TG{check if indeed proprotional, add regression slopes}.

On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
slightly lower than Dask's in particular as the number of workers increase.
Within Dask, Delayed and Futures have a higher overhead than Bags. However,
overhead differences are compensated by an increase in data transfer time,
as a reduced overhead increases the concurrency between data transfers.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_worker}
    \end{subfigure}
    \caption{125 blocks, 10 iterations, \SI{4}{\second} delay, 8
    threads/worker}\label{fig:inc_worker}
\end{figure}

@@

In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
For 30 blocks the Dask Bag API has a much lower overhead time. This is because Bag
was only using one thread per block in comparison to Delayed and Futures that were
offloading block calculations on multiple threads of the same worker. Once again,
the data transfer time reduces with more blocks however the overhead time increases
by a similar amount. This is not observed for 30 blocks as the workers are not used
at full capacity, i.e., some threads are idle. Finally, the variability of the
overhead increases with the number of blocks, which explains the makespan variability
mentioned previously.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_block.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_block.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_block}
    \end{subfigure}
    \caption{10 iterations, \SI{4}{\second} delay, 8 workers, 8
    threads/worker}\label{fig:inc_block}
\end{figure}

%% Iterations
\subsection{Incrementation: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application while
varying the number of iterations. Overall,  Spark and Dask APIs are once
again equivalent, although Dask Delayed and Futures are slightly faster
than Bags and, Spark for 1 and 10 few iterations, and Futures are faster than
Delayed, Bags and Spark for 100 iterations. Differences remain minor
though.

In Figure~\ref{fig:inc_tt_itr}, the total execution time breakdown is
shown. We observe the good scalability of all the engines with the number of
iterations.
% TODO

% We think this is because Dask Delayed often
% schedule a block on a different thread of the same worker which causes small delays
% every time due to inter-thread data communication. This is good when there are fewer
% tasks as is make the IO between blocks go out of sync hence lowering the IO
% bottleneck however as the number of tasks increases those small delays become
% significant.

%  The Dask Futures API seems to outperform all the other APIs but this is
% most likely because the tasks are not interdependent thus this application benefits
% from the less optimal but faster scheduling brought by Dask Futures.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_itr}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_itr}
    \end{subfigure}
    \caption{125 blocks, \SI{4}{\second} delay, 8 workers, 8 threads/worker}
\end{figure}

@@

Figure~\ref{fig:inc_tt_sleep} shows the total execution time breakdown. On
one hand, Spark has the smallest overhead. As previously observed,
variations in overhead time are almost exactly compensated by variations in data
transfer time.

% \begin{table}[!t]
%     \caption{Function time for the sleep experiment}
%     \begin{subtable}[b]{\columnwidth}
%         \renewcommand{\arraystretch}{1.3}
%         \caption{Overhead time in second}\label{tab:inc_sleep_overhead}
%         \centering
%         \begin{tabular}{lllll}
%         \hline
%                      & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
%         Spark        & 14992  & 17076  & 16069   & 9769    \\
%         Dask.Bag     & 20690  & 21388  & 17759   & 18720   \\
%         Dask.Delayed & 27481  & 26199  & 22080   & 24624   \\
%         Dask.Futures & 27069  & 27984  & 23538   & 24762   \\ \hline
%         \end{tabular}
%     \end{subtable}
%     \vskip 0.2cm
%     \begin{subtable}[b]{\columnwidth}
%         \renewcommand{\arraystretch}{1.3}
%         \caption{IO time in second}\label{tab:inc_sleep_io}
%         \centering
%         \begin{tabular}{lllll}
%         \hline
%                      & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
%         Spark        & 63622  & 57144  & 51070   & 57078   \\
%         Dask.Bag     & 54630  & 52312  & 53743   & 53612   \\
%         Dask.Delayed & 45239  & 44142  & 45133   & 46198   \\
%         Dask.Futures & 46537  & 43161  & 44749   & 45846   \\ \hline
%         \end{tabular}
%     \end{subtable}
%     \vspace{-3mm}
%  \end{table}

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_sleep}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_sleep}
    \end{subfigure}
    \caption{125 blocks, 10 iterations, 8 workers, 8 threads/worker}
\end{figure}

@@

Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine and API.
Gantt charts are structured in batches of up to 64 read-compute-write concurrent
sequences. File reads in the first batch are much longer than in the following ones:
this is due to the high synchronization of data transfers that leads to a high
saturation of the shared file system. We also note that overhead, represented in
white, is concentrated around the data transfer tasks \TG{find out why}, and the
computing tasks that run concurrently with data transfers. 
% The frameworks have similar makespan however the time spend for each type of
% task differs considerably (see Table~\ref{tab:inc_base}). Spark tends to spend more
% time than other frameworks for IO however it has a much lower overhead. Dask Delayed
% and Futures have the lowest IO time but a much higher overhead. Dask Bag stands in
% the middle for both IO and overhead. Compute time is approximately the same for all
% frameworks. Apart from the first IO section, there is a large amount of ovehead
% everytime time IO is performed. This is true for all frameworks.

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/spark_inc_gantt.png}
        \caption{Spark}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/bag_inc_gantt.png}%
        \caption{Dask Bag}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/delayed_inc_gantt.png}%
        \caption{Dask Delayed}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/futures_inc_gantt.png}%
        \caption{Dask Futures}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{Incrementation: Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second}
delay, 8 workers, 8 threads/worker}\label{fig:inc_gantt}
\end{figure*}

% \begin{table}[!b]
%     \renewcommand{\arraystretch}{1.3}
%     \caption{Distribution of the execution time in second: 125 blocks, 4 sec.\ sleep
%     delay, 10 iterations, and 8 workers}\label{tab:inc_base}
%     \centering
%     \begin{tabular*}{\columnwidth}{llllll}
%     \hline
%                  & Read  & Compute & Write & Overhead & Total \\ \hline
%     Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
%     Dask Bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
%     Dask Delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
%     Dask Futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
%     \end{tabular*}
%  \end{table}

%%% HISTOGRAM %%%
\subsection{Histogram: Number of workers}
Figure~\ref{fig:histo_ms_worker} shows the makespan of the application for various
amounts of workers. Spark is tremendously faster than Dask APIs. The difference
narrows as the number of workers increases. Between Dask APIs there is no substantial
difference, however, Bag is slightly faster on average.

From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is shown.
The computation time is significantly larger in Dask than Spark. This is potentially
due to Python's GIL preventing Dask to parallelize the computation on multiple threads.
Overall, the IO and overhead are comparable for all engines.

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_worker.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_worker.png}%
        \caption{Histogram total time}\label{fig:histo_tt_worker}
    \end{subfigure}
    \caption{125 blocks, 8 threads/worker}
\end{figure}

\subsection{Histogram: Number of blocks}
Figure~\ref{fig:histo_ms_block} shows the makespan for different block size.
Again, Spark is significantly faster then Dask engines and Bag is slightly faster
than Delayed. Overall, the engines do not react to change in block size.

@@

Spark has a much lower compute time which makes it significantly faster than Dask
engines. Overall, for all engines, lower block size results in lower IO time but
increases the overhead time. Again, this is due to the overhead desynchronizing task
thus reducing the NFS bottleneck.
%TODO

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_block.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_block.png}%
        \caption{Histogram total time}\label{fig:histo_tt_block}
    \end{subfigure}
    \caption{8 workers, 8 threads/worker}
\end{figure}
\TG{We decided to omit Dask Futures as it does not have any value over Dask
Delayed in this application.}

@@

The reduce is done differently on the engines. While Spark and Delayed compute all
the reduce tasks on one process (thread) Spark computes it on the driver process. On
the other hand, Bag computes the reduce tasks on different workers and sends the
information to the driver process afterward.

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/spark_histo_gantt.png}
        \caption{Spark}\label{fig:histo_spark_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/bag_histo_gantt.png}%
        \caption{Dask Bag}\label{fig:histo_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/delayed_histo_gantt.png}%
        \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
    \end{subfigure}
    \caption{Histogram: Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second}
delay, 8 workers, 8 threads/worker}\label{fig:histo_gantt}
\end{figure}

%%% HISTOGRAM (NumPy) %%%
\subsection{Histogram (NumPy): Number of workers}
Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount of workers.
Overall, the makespan for all engines does not change considerably with an increase
in the number of workers. This is because the engines do not scale well.

On figure~\ref{fig:histo_np_ms_worker}, there is no substantial difference between
 all engines, although, Delayed is slightly faster.

From Figure~\ref{fig:histo_np_tt_worker}, the total execution time spent in each
function is shown. The overhead and read time increase proportionally to the number
of workers thus increasing the total time. This is because (1) most workers read at
the same time due to a low computation-IO ratio and (2) \MD{Determine why the
overhead is so large compared to the pure Python implementation}

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_np_worker.png}%
        \caption{Histogram (NumPy) makespan}\label{fig:histo_np_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_np_worker.png}%
        \caption{Histogram (NumPy) total time}\label{fig:histo_np_tt_worker}
    \end{subfigure}
    \caption{125 blocks, 8 threads/worker}
\end{figure}

\subsection{Histogram (NumPy): Number of blocks}
Figure~\ref{fig:histo_np_ms_block} shows the makespan of the application for
different block size. Note that the workers can process up to 64 tasks in parallel
thus they are not used at full capacity with 30 blocks. Overall, there is no
substantial difference for all engines and block size.

In figure~\ref{fig:histo_np_tt_block}, shows the total time spent in each function.
Again, the overhead for 30 blocks is erroneous due to a lower amount of task than
available threads which result in higher overhead for thread-based engines.
Considering the overhead error and that at 30 blocks the workers only use half their
core, the total time is substantially the same for all block size. This is due to
lower block sizes having a lower IO time, however, it is balanced by the resulting
increase in overhead.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_np_block.png}%
        \caption{Histogram (NumPy) makespan}\label{fig:histo_np_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_np_block.png}%
        \caption{Histogram (NumPy) total time}\label{fig:histo_np_tt_block}
    \end{subfigure}
    \caption{8 workers, 8 threads/worker}
\end{figure}

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{times}
\usepackage[binary-units=true]{siunitx}
\usepackage{latexsym}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{dblfloatfix}

@@

While Python multithreading is limited by the
\href{https://docs.python.org/3/glossary.html#term-gil}{Global Interpreter Lock (GIL)}
as a result of CPython's non thread-safe memory management, Dask attempts to
bypass the GIL whenever possible and use multithreading to provide parallelism.
With the exception of Dask Bag, all APIs, by default, use the local multithreaded
scheduler. Dask Bag instead relies on the local multiprocessing scheduler.

@@

\begin{figure}[!b]
    \centering
    \includegraphics[width=2cm, height=\columnwidth,
    angle=-90]{images/incrementation-task-graph.png}
    \caption{Task graph for Incrementation with 5 iterations and 3 BigBrain blocks.
    Circles represent the incrementation and sleep function while rectangles
    represent stages of the blocks.}\label{fig:tg-inc}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=2cm, height=\columnwidth,
    angle=-90]{images/histogram-task-graph.png}
    \caption{Task graph for Histogram with 3 BigBrain blocks. Circles represent the
    functions while rectangles represent stages of the blocks.
    }\label{fig:tg-histo}
\end{figure}

@@

While Incrementation and Histogram were implemented directly in Python, BIDS App
example requires an external library and is distributed as a Docker container image
(bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image,
given its popularity for HPC and reproducibility, using
\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}.
The image was preloaded on the nfs and singularity version 3.2.1\-1.el7 was installed
on all instances. Due to a bug in the conversion from Docker to Singularity, we had to
export the path for fsl.

@@

For each run, we measured the application makespan as well as the cumulative 
data read, compute, data write, and engine overhead across all application tasks. 
We measured the overhead by

% \TG{\ldots, maybe add a figure}. \MD{Done but might need
% adjustment depending on how the overhead is calculate (idle vs wasted)}

% \TG{explain that number of iterations may impact futures.}

The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
seconds sleep delay.

@@

%%% INCREMENTATION %%%
%% worker
\subsection{Incrementation: Number of workers}
Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation application
for different numbers of workers and engines. The bars show the average
makespan over 3 repetitions while the error bars are the standard deviation. Overall,
there is no substantial makespan difference between the engines. Dask seems to have a
slight advantage over Spark, \SI{83.61}{\second} on average,
with Delayed and Futures being slightly better than Bag.

For all engines, the makespan is far from decreasing linearly with the
number of workers. Note that each worker has 8 CPUs. The makespan
even increases between 4 and 8 workers. This is due to the high impact of
data transfers and engine overhead on the application.

Figure~\ref{fig:inc_tt_worker} shows the total execution time of the
Incrementation application, broken down into data transfer (read and
write), compute (sleep), and overhead time. As expected, the computing time
stays similar when the number of workers increases. However, the data
transfer time and overhead increase proportionally to the number of
workers with regression lines:

\hspace{1cm}\(\hat{y}=2336.9499\cdot{x} + 325.00574\) (Spark)

\hspace{1cm}\(\hat{y}=2650.13505\cdot{x} + 1585.37218\) (Bag)

\hspace{1cm}\(\hat{y}=3250.53406\cdot{x} + 1769.50351\) (Delayed)

\hspace{1cm}\(\hat{y}=3570.39758\cdot{x} + 1062.23031\) (Futures)

On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
slightly lower than Dask's in particular as the number of workers increase.
Within Dask, Delayed and Futures have a higher overhead than Bag. However,
overhead differences are compensated by an increase in data transfer time,
as a reduced overhead increases the concurrency between data transfers.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_worker}
    \end{subfigure}
    \caption{125 blocks, 10 iterations, \SI{4}{\second} delay, 8
    CPUs/worker}\label{fig:inc_worker}
\end{figure}

@@

In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
For 30 blocks, Delayed and Futres have much higher overhead. This is because they are
thread-based and use multiple threads to compute a block which affect the calculation
of the overhead. Once again, the data transfer time reduces with more blocks however
the overhead time increases by a similar amount. This is not observed for 30 blocks
as the workers are not used at full capacity, i.e., some threads are idle. Finally,
the variability of the overhead increases with the number of blocks, which explains
the makespan variability mentioned previously.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_block.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_block.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_block}
    \end{subfigure}
    \caption{10 iterations, \SI{4}{\second} delay, 8 workers, 8
    CPUs/worker}\label{fig:inc_block}
\end{figure}

%% Iterations
\subsection{Incrementation: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application while
varying the number of iterations. Overall,  Spark and Dask APIs are once
again equivalent, although Delayed and Futures are slightly faster
than Bag and, Spark for 1 and 10 iterations, and Futures are faster than
Delayed, Bag and Spark for 100 iterations. Differences remain minor
though.

In Figure~\ref{fig:inc_tt_itr}, the total execution time breakdown is
shown. We observe the good scalability of all the engines with the number of
iterations.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_itr}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_itr}
    \end{subfigure}
    \caption{125 blocks, \SI{4}{\second} delay, 8 workers, 8 CPUs/worker}
\end{figure}

@@

Figure~\ref{fig:inc_tt_sleep} shows the total execution time breakdown. Spark has the
smallest overhead. As previously observed, variations in overhead time are almost
exactly compensated by variations in data transfer time.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_sleep}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_sleep}
    \end{subfigure}
    \caption{125 blocks, 10 iterations, 8 workers, 8 CPUs/worker}
\end{figure}

@@

Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine. Gantt
charts are structured in batches of up to 64 read-compute-write concurrent sequences.
Files read in the first batch are much longer than the ones in the following batches:
this is due to the high synchronization of data transfers that leads to a high
saturation of the shared file system. We also note that overhead, represented in
white, is concentrated around the data transfer tasks, and the
computing tasks that run concurrently with data transfers.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/spark_inc_gantt.png}
        \caption{Spark}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/bag_inc_gantt.png}%
        \caption{Dask Bag}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/delayed_inc_gantt.png}%
        \caption{Dask Delayed}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/futures_inc_gantt.png}%
        \caption{Dask Futures}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{Incrementation: 125 blocks, 10 iterations, \SI{4}{\second} delay, 8 workers,
8 CPUs/worker}\label{fig:inc_gantt}
\end{figure}

%%% HISTOGRAM %%%
\subsection{Histogram: Number of workers}
Figure~\ref{fig:histo_ms_worker} shows the makespan of the application for various
amounts of workers. Spark is tremendously faster than Dask APIs. The difference
narrows as the number of workers increases but stays significant. Between Dask APIs
there is no substantial difference, however, Bag is slightly faster on average.

From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is
shown. The computation time is significantly larger for Dask than for Spark. This is
potentially due to Python's GIL preventing Dask to parallelize the computation on
multiple threads. Overall, the IO and overhead are comparable for all engines.

@@

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_worker.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_worker.png}%
        \caption{Histogram total time}\label{fig:histo_tt_worker}
    \end{subfigure}
    \caption{125 blocks, 8 CPUs/worker}
\end{figure}

\subsection{Histogram: Number of blocks}
Figure~\ref{fig:histo_ms_block} shows the makespan for different block size. Again,
Spark is not compared at 30 blocks due to its \SI{2}{\giga\byte} task limitation.
Once more, Spark is significantly faster then Dask engines and Bag is slightly faster
than Delayed. Overall, the engines do not react to change in block size.

@@

Spark has a much lower compute time which makes it significantly faster than Dask
engines. Overall, for all engines, lower block size results in lower IO time but
increases the overhead time. Again, this is due to large overhead desynchronizing task
thus reducing the NFS bottleneck.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_block.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_block.png}%
        \caption{Histogram total time}\label{fig:histo_tt_block}
    \end{subfigure}
    \caption{8 workers, 8 CPUs/worker}
\end{figure}
% \TG{We decided to omit Dask Futures as it does not have any value over Dask
% Delayed in this application.}

@@

The reduce is done differently for the engines. Spark and Delayed compute all the
reduce tasks on one process (thread), however, Spark only computes it on the driver
process. On the other hand, Bag computes the reduce tasks on different workers and
sends the information to the driver process afterward.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/spark_histo_gantt.png}
        \caption{Spark}\label{fig:histo_spark_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/bag_histo_gantt.png}%
        \caption{Dask Bag}\label{fig:histo_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/delayed_histo_gantt.png}%
        \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
    \end{subfigure}
    \caption{Histogram: 125 blocks, 8 workers, 8
    CPUs/worker}\label{fig:histo_gantt}
\end{figure}

%%% HISTOGRAM (NumPy) %%%
\subsection{Histogram (NumPy): Number of workers}
Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount of workers.
Overall, the makespan for all engines does not change considerably with an increase
in the number of workers. This is because the engines do not scale well. There is no
substantial difference between all engines, although, Delayed is slightly faster.

From Figure~\ref{fig:histo_np_tt_worker}, the total execution time breakdown is
shown. The overhead and read time increase proportionally to the number of workers
thus increasing the total time. This is because (1) most workers read at the same
time due to a low computation-IO ratio and (2) there is more intra-worker and
inter-worker communication.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_np_worker.png}%
        \caption{Histogram (NumPy) makespan}\label{fig:histo_np_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_np_worker.png}%
        \caption{Histogram (NumPy) total time}\label{fig:histo_np_tt_worker}
    \end{subfigure}
    \caption{125 blocks, 8 CPUs/worker}
\end{figure}

\subsection{Histogram (NumPy): Number of blocks}
Figure~\ref{fig:histo_np_ms_block} shows the makespan of the application for
different block size. Again, Spark is not compared at 30 blocks due to its
\SI{2}{\giga\byte} task limitation. Overall, there is no substantial difference for
all engines and block sizes.

In figure~\ref{fig:histo_np_tt_block}, shows the total time spent in each function.
Once more, since Delayed is thread-based its overhead at 30 blocks is seems much
higher that it really is. Considering the overhead error and that at 30 blocks the
workers only use half their core, the total time is substantially the same for all
block size. This is due to lower block sizes having a lower IO time, however, it
balances out by the resulting increase in overhead.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_np_block.png}%
        \caption{Histogram (NumPy) makespan}\label{fig:histo_np_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_np_block.png}%
        \caption{Histogram (NumPy) total time}\label{fig:histo_np_tt_block}
    \end{subfigure}
    \caption{8 workers, 8 CPUs/worker}
\end{figure}

\subsection{Histogram (NumPy): Gantt chart}
Figure~\ref{fig:histo_np_gantt} shows the Gantt chart of the baseline for the
Histogram (NumPy) experiment. The overhead for Spark and Bag is mostly located
between a read and compute task or compute and read task. On the other hand, Delayed
overhead is dispersed between all types of tasks. This causes Delayed to overlap more
compute and read tasks simultaneously among the different workers.

In figure~\ref{fig:histo_np_gantt}, read tasks are initially similar for all engines.
However, when computation starts the subsequent read tasks are much shorter for
Delayed. This is due the desynchronization of the task reducing the IO bottleneck.
